Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-03-23 ms_pre_post.Rmd
Remaining edits to manuscript

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance
  and associated polyandry
author: ''
date: "19 March 2017"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

In all simulations, we recorded mean values of $P_{a}$, $M_{a}$, and $F_{a}$ in each generation, and we present these values over generations to infer selection on phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$). Each combination of parameter values simulated was replicated $40$ times, and grand mean values and standard errors of means are calculated in each generation across replicates. These analyses allowed us to infer how allele values changed over generations in response to costs, but also in response to the changing values of other alleles and therefore potential evolutionary feedbacks among reproductive strategies. For all replicates, we set the maximum number of generations to 40000, which exploratory simulations and previous modelling [@Duthie] showed to be sufficient for inferring long-term dynamics of mean allele values and therefore selection on phenotypes.  For all replicates, we set $\rho=5$ immigrants, which produced a range of kin and non-kin in each generation allowing females to express inbreeding strategies, and $n=8$ offspring, which was sufficient to keep populations consistently at carrying capacities and avoid population extinction. Values of $K_{f}$ and $K_{m}$ were both set to 100 because previous modelling showed that populations of this size are small enough that mate encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but not so small that selection is typically overwhelmed by drift [@Duthie2016a].

@@

When post-copulatory inbreeding avoidance was allowed to evolve, mean $P_{a}$ values became substantially higher than in comparable simulations where $F_{a}$ values were fixed to zero and post-copulatory inbreeding avoidance could not evolve (Figure 1B,D,F,H). Allowing evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance therefore facilitated evolution of polyandry to the degree that most females mated multiply given low costs of polyandry ($c_{P} < 0.005$; e.g., Figure 2A,B). Hence, given a low cost of polyandry, $P_{a}$ allele values increased from zero and persisted at low positive values (Figure 1B,D). In contrast, higher costs of polyandry ($c_{P} \geq 0.005$) resulted in initial increases of $P_{a}$ allele values from zero, but then became slightly negative over generations. Trajectories of allele values in individual simulations were typically highly stochastic, but also consistent in long-term direction (Supporting Information pp. S5-S12). Overall, this implies that, even given strong inbreeding depression in offspring viability, inbreeding avoidance might generally be a weak force driving polyandry evolution.

@@

When pre-copulatory inbreeding avoidance was costly, allowing evolution of cost free post-copulatory inbreeding avoidance greatly facilitated evolution of polyandry (Figure 3A versus 3B). However when post-copulatory inbreeding avoidance was costly, allowing evolution of cost free pre-copulatory inbreeding avoidance caused $P_{a}$ alleles to decrease to very negative values, reducing expression of polyandry (Figure 3C versus 3D; polyandry was cost free in all these simulations).

@@

Conversely, when pre-copulatory inbreeding allele ($M_{a}$) values were fixed for adaptive inbreeding avoidance, mean $F_{a}$ allele values did not consistently become negative over generations (Figure 4C,D). Fixation of pre-copulatory inbreeding avoidance therefore prevented evolution of post-copulatory inbreeding avoidance. However, in this case, mean $P_{a}$ allele values consistently decreased over generations, reflecting selection against polyandry regardless whether or not post-copulatory inbreeding avoidance was costly (Figure 4C,D). Consequently, when $c_{F}=0$, $F_{a}$ allele values had no effect because females were almost exclusively monandrous, resulting in high drift of $F_{a}$ values among replicates (illustrated by wide standard errors, Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs. This lack of selection for post-copulatory inbreeding avoidance was driven by a lack of polyandry, and therefore an inability of females to bias fertilisation among multiple mates. When pre-copulatory inbreeding avoidance and polyandry were both fixed ($M_{a}=-10$ and $P_{a}=1$), $F_{a}$ allele values evolved to similarly negative means as $M_{a}$ allele values in Figure 4A,B (see Supporting Information).

@@

Evolution of reproductive strategies cannot be presumed to occur in isolation from each other in real-world populations, and is instead likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be defined by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.

@@

The opportunity to adjust inbreeding has been widely considered as a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. We found that when post-copulatory inbreeding avoidance could evolve, selection for and resulting evolution of polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but our model clarifies this verbal hypothesis and therefore has widespread implications for future studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and inbreeding avoidance through post-copulatory mechanisms can also evolve. Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g., @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. Recent studies have shown how females can successfully bias fertilisation after mating has occurred. For example, when female guppies (*Poecilia reticulata*) were artificially inseminated with equal quantities of sperm from full-siblings and unrelated males, more eggs were fertilised by unrelated males because the velocities of full siblings' sperm were reduced by females' ovarian fluids [@Gasparini2011]. In black field crickets (*Teleogryllus commodus*), females attempt to remove the spermatophores of unwanted males after copulation, and are capable of controlling sperm transfer to spermatheca after copulation occurs [@Bussiere2006; @Tuni2013]. Our results suggest that evolution of such post-copulatory inbreeding avoidance might increase selection for further polyandry, especially if pre-copulatory inbreeding avoidance is costly (Figure 1).

@@

Evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance occurred in our model, but was affected by the joint evolution of polyandry and the existence of asymmetric costs. One cost free strategy of inbreeding avoidance precluded another more costly strategy from persisting in a focal population by modifying its effect on inbreeding avoidance (Figure 3). Consequently, long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given sufficient cost asymmetry. Direct costs therefore strongly affected the evolution of the costly phenotype(s), meaning that quantifying costs is necessary for predicting evolution of phenotypes associated with inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which are also relevant for polyandry, and include costs of elevated risks of predation or disease transmission, and costs incurred through time or energy expenditure. Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Here we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs thereby reflect empirical observations in some populations and are therefore biologically realistic, other types of costs have also been observed and could therefore be considered in future models. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Different types of costs might affect selection both directly and indirectly through interactions between reproductive strategies, thereby affecting evolution of entire reproductive systems. Evolution of individual phenotypes therefore needs to be considered in the broader context of potentially interacting and phenotypes and their costs.

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also naturally depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and polyandry is uncommon [@Lihoreau2007; @Metzger2010; @Metzger2010a], evolution of post-copulatory inbreeding avoidance is highly unlikely.

@@

@@

This work was funded by a European Research Council Starting Grant to JMR.

@@

![Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black) from simulations where post-copulatory inbreeding strategy is (A, C, E, and G) fixed to zero (i.e., random fertilisation) or (B, D, F, H) allowed to evolve freely. Costs of polyandry (c_p) increase across rows from zero (A and B) to 0.01 (G and H). Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate inbreeding avoidance or tendency for monandry, and positive values indicate inbreeding preference or tendency for polyandry. Dotted lines demarcate mean allele values of zero.](figures/poly_costs.png)

@@

![Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black) when (A and B) costly pre-copulatory inbreeding strategy ($c_{M} = 0.02$) can evolve and post-copulatory inbreeding strategy is (A) fixed for random fertilisation or (B) can also evolve, and when (C and D) costly post-copulatory inbreeding avoidance ($c_{F} = 0.02$) can evolve and pre-copulatory inbreeding strategy is (A) fixed for random mating or (B) can also evolve. Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry. In all panels, polyandry is cost free.](figures/rel_costs.png)

![Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black), given (A and B) fixed polyandry and post-copulatory inbreeding avoidance and (C and D) fixed post-copulatory inbreeding avoidance where the evolving inbreeding strategy is cost-free (A, $c_{M} = 0$; C, $c_{F} = 0$) or costly (B, $c_{M}=0.02$; D, $c_{F}=0.02$). Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values reflect strategies of inbreeding avoidance or tendency for monandry, and positive values reflect strategies of inbreeding preference or tendency for polyandry.](figures/fixed_adapt.png)

>>>>

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance
  and associated polyandry
author: ''
date: "23 March 2017"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

In all simulations, we recorded mean values of $P_{a}$, $M_{a}$, and $F_{a}$ in each generation and present these values over generations to infer selection on phenotypes ($P_{p}$, $M_{p}$, and $F_{p}$). Each combination of parameter values simulated was replicated $40$ times, and grand mean values and standard errors of means are calculated in each generation across replicates. These analyses allowed us to infer how allele values changed over generations in response to costs, but also in response to the changing values of other alleles and therefore potential evolutionary feedbacks among reproductive strategies. For all replicates, we set the maximum number of generations to 40000, which exploratory simulations and previous modelling [@Duthie] showed to be sufficient for inferring long-term dynamics of mean allele values and therefore selection on phenotypes.  For all replicates, we set $\rho=5$ immigrants, which produced a range of kin and non-kin in each generation allowing females to express inbreeding strategies, and $n=8$ offspring, which was sufficient to keep populations consistently at carrying capacities and avoid population extinction. Values of $K_{f}$ and $K_{m}$ were both set to 100 because previous modelling showed that populations of this size are small enough that mate encounters between kin occur with sufficient frequency for selection on inbreeding strategy, but not so small that selection is typically overwhelmed by drift [@Duthie2016a].

@@

When post-copulatory inbreeding avoidance was allowed to evolve, mean $P_{a}$ values became substantially higher than in comparable simulations where $F_{a}$ values were fixed to zero and post-copulatory inbreeding avoidance could not evolve (Figure 1B,D,F,H). Allowing evolution of post-copulatory inbreeding avoidance alongside pre-copulatory inbreeding avoidance therefore facilitated evolution of polyandry to the degree that most females mated multiply given low costs of polyandry ($c_{P} < 0.005$; e.g., Figure 2A,B). Here, $P_{a}$ allele values increased from zero and persisted at low positive values (Figure 1B,D). Given higher costs of polyandry ($c_{P} \geq 0.005$), $P_{a}$ allele values still initially increased from zero, but then became slightly negative over generations (Figure 1F,H). Trajectories of allele values in individual simulations were typically highly stochastic, but were consistent in their long-term direction (Supporting Information pp. S5-S12). Overall, this implies that, even given strong inbreeding depression in offspring viability, inbreeding avoidance might generally be a weak force driving polyandry evolution.

@@

When pre-copulatory inbreeding avoidance was costly, allowing evolution of cost free post-copulatory inbreeding avoidance greatly facilitated evolution of polyandry (Figure 3A versus 3B). However when post-copulatory inbreeding avoidance was costly, allowing evolution of cost free pre-copulatory inbreeding avoidance caused $P_{a}$ alleles to decrease to very negative values, reducing expression of polyandry (Figure 3C versus 3D; polyandry was cost free in all these simulations). Results for all possible cost combinations of $0$ and $0.02$ are provided in Supporting Information (pp. S5-S12).

@@

Conversely, when pre-copulatory inbreeding allele ($M_{a}$) values were fixed for adaptive inbreeding avoidance, mean $F_{a}$ allele values did not consistently become negative over generations (Figure 4C,D). Fixation of pre-copulatory inbreeding avoidance therefore prevented evolution of post-copulatory inbreeding avoidance. However, in this case, mean $P_{a}$ allele values consistently decreased over generations, reflecting selection against polyandry regardless whether or not post-copulatory inbreeding avoidance was costly (Figure 4C,D). Consequently, when $c_{F}=0$, $F_{a}$ allele values had no effect because females were almost exclusively monandrous, resulting in high drift of $F_{a}$ values among replicates (illustrated by wide standard errors, Figure 4C). However, when $c_{F}=0.02$, $F_{a}$ values remained near zero to minimise direct costs. This lack of selection for post-copulatory inbreeding avoidance was driven by a lack of polyandry, and therefore an inability of females to bias fertilisation among multiple mates. When pre-copulatory inbreeding avoidance and polyandry were both fixed ($M_{a}=-10$ and $P_{a}=1$), $F_{a}$ allele values evolved to similarly negative means as $M_{a}$ allele values in Figure 4A,B (see Supporting Information p. S4).

@@

Evolution of different reproductive strategies cannot be presumed to occur in isolation from each other in real-world populations. Rather, such evolution is likely to involve multiple interacting phenotypes resulting in the potential for feedback and degeneracy (i.e., functional redundance) among phenotypes under selection. For example, strong inbreeding depression in offspring fitness is widely hypothesised to drive selection for inbreeding avoidance, which could consequently drive selection for polyandry [e.g., @Zeh1997; @Jennions2000; @Tregenza2002]. Evolution of inbreeding avoidance can be driven by selection on both pre-copulatory and post-copulatory mechanisms, so simultaneous evolution of each phenotype is likely to be defined by degeneracy, in addition to any direct costs on phenotypes. We used individual-based modelling to explore evolution of pre-copulatory and post-copulatory inbreeding avoidance and polyandry given (1) hard constraints on evolution of post-copulatory inbreeding avoidance, (2) asymmetric costs of pre-copulatory and post-copulatory inbreeding avoidance, and (3) evolution of one inbreeding avoidance phenotype given the pre-existence the other. We thereby highlight fundamental but theoretically under-developed relationships between evolution of female mating frequency and pre-copulatory versus post-copulatory inbreeding avoidance.

@@

The opportunity to adjust inbreeding has been widely suggested to be a driver of adaptive evolution of polyandry [@Tregenza2002; @Foerster2003; @Akcay2007; @Varian-Ramos2012; @Kingma2013; @Lehtonen2015; @Reid2014]. We found that when post-copulatory inbreeding avoidance could evolve, selection for and resulting evolution of polyandry was greatly strengthened (Figure 1). The proposition that polyandry might facilitate cryptic female choice among males of varying compatibility is not new [e.g., @Zeh1997], but our model clarifies this verbal hypothesis and therefore has widespread implications for future studies of evolution of polyandry.

We predict evolution of polyandry in populations where inbreeding depression is severe and inbreeding avoidance through post-copulatory mechanisms can also evolve, especially if pre-copulatory inbreeding avoidance is costly (Figure 1). Post-copulatory mechanisms of inbreeding avoidance have been observed under these conditions in experimental systems across diverse taxa [e.g., @Pizzari2004; @Firman2008; @Bretman2009; @Gasparini2011; @Tuni2013; @Firman2015]. Recent studies have shown how females can successfully bias fertilisation after mating has occurred. For example, when female guppies (*Poecilia reticulata*) were artificially inseminated with equal quantities of sperm from full-siblings and unrelated males, more eggs were fertilised by unrelated males because the velocities of full siblings' sperm were reduced by females' ovarian fluids [@Gasparini2011]. In black field crickets (*Teleogryllus commodus*), females attempt to remove the spermatophores of unwanted males after copulation, and are capable of controlling sperm transfer to spermatheca after copulation occurs [@Bussiere2006; @Tuni2013].

@@

Evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance occurred in our model, but was affected by the joint evolution of polyandry and the existence of asymmetric costs. One cost free strategy of inbreeding avoidance precluded another more costly strategy from persisting in a focal population by modifying its effect on inbreeding avoidance (Figure 3). Consequently, long-term persistence of both pre-copulatory and post-copulatory inbreeding avoidance might not be expected in populations given sufficient cost asymmetry. Since direct costs strongly affected the evolution of the focal phenotype(s), quantifying costs is necessary for predicting evolution of inbreeding strategies. Quantifying direct costs of phenotypes associated with polyandry and mate choice is empirically challenging [@Pomiankowski1987; @Kokko2003]. @Pomiankowski1987 categorised four types of costs that are relevant to mate choice, which are also relevant for polyandry, and include costs of elevated risks of predation or disease transmission, and costs incurred through time or energy expenditure. Another type of cost relevant for polyandry and inbreeding strategy is risk of male harm caused by sexual conflict [e.g., @Arnqvist2005a; @Parker2006]. We assumed that males were passive in mating encounters with females, but inbreeding theory predicts that males should benefit from a higher tolerance of inbreeding than females, leading to sexual conflict over inbreeding [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. Here we interpreted a cost of polyandy as an elevated risk of predation as might be incurred while searching for or courting mates [e.g., @Rowe1988; @Rowe1994], a cost of pre-copulatory inbreeding strategy as a lost of time [i.e., risk of not finding a mate in time due to choosiness; e.g., @Kokko2013], and a cost of post-copulatory inbreeding strategy as depletion of energy. While these costs thereby reflect empirical observations in some populations and are therefore biologically realistic, other types of costs have also been observed and could therefore be considered in future models. For example, increased risk of disease transmission has been observed for highly polyandrous females [@Roberts2015a], but such a cost would more realistically be applied to the number of realised mates a female has rather than her tendency for engaging in polyandry. Different types of costs might affect selection both directly and indirectly through interactions between reproductive strategies, thereby affecting evolution of entire reproductive systems. Evolution of individual phenotypes therefore needs to be considered in the broader context of potentially interacting and phenotypes and their costs.

In our model, existence of adaptive pre-copulatory inbreeding avoidance precluded evolution of polyandry and, in turn, evolution of post-copulatory inbreeding avoidance (Figure 3C,D). However, existence of adaptive post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance (Figure 3A,B). In wild populations, it is unlikely that pre-copulatory and post-copulatory inbreeding avoidance will evolve simultaneously from an ancestral population in which females mate and assign paternity randomly. Rather, for example, initial evolution of pre-copulatory inbreeding avoidance is much more likely to evolve in the absence of evolving post-copulatory inbreeding avoidance due to constraints, or in the presence of already evolved post-copulatory inbreeding avoidance. When framing hypotheses for existence of post-copulatory inbreeding avoidance and polyandry, it might therefore be necessary to consider whether or not inbreeding avoidance already occurs through pre-copulatory mate choice. Additionally, the opportunity for post-copulatory inbreeding avoidance will also depend on the degree to which females are polyandrous. For species in which pre-copulatory inbreeding avoidance occurs and polyandry is uncommon [@Lihoreau2007; @Metzger2010; @Metzger2010a], evolution of post-copulatory inbreeding avoidance is highly unlikely even if such a strategy incurs little direct cost.

@@

<!--- TEMPORARILY REMOVED FOR DOUBLE-BLIND PEER REVIEW

@@

This work was funded by a European Research Council Starting Grant to JMR. Computer simulations were performed using the Maxwell Computing Cluster at the University of Aberdeen.

@@

![Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black) from simulations where post-copulatory inbreeding strategy is (A, C, E, and G) fixed to zero (i.e., random fertilisation) or (B, D, F, and H) allowed to evolve freely. Costs of polyandry (c_p) increase across rows from zero (A and B) to 0.01 (G and H). Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate inbreeding avoidance or tendency for monandry, and positive values indicate inbreeding preference or tendency for polyandry. Dotted lines demarcate mean allele values of zero.](figures/poly_costs.png)

@@

![Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black) when (A and B) costly pre-copulatory inbreeding strategy ($c_{M} = 0.02$) can evolve and post-copulatory inbreeding strategy is (A) fixed for random fertilisation or (B) can also evolve, and when (C and D) costly post-copulatory inbreeding avoidance ($c_{F} = 0.02$) can evolve and pre-copulatory inbreeding strategy is (C) fixed for random mating or (D) can also evolve. Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate strategies of inbreeding avoidance or tendency for monandry, and positive values indicate strategies of inbreeding preference or tendency for polyandry. In all panels, polyandry is cost free.](figures/rel_costs.png)

![Mean allele values underlying tendency for polyandry (red), pre-copulatory inbreeding strategy (blue), and post-copulatory inbreeding strategy (black), given (A and B) fixed polyandry and post-copulatory inbreeding avoidance and (C and D) fixed post-copulatory inbreeding avoidance where the evolving inbreeding strategy is cost-free (A, $c_{M} = 0$; C, $c_{F} = 0$) or costly (B, $c_{M}=0.02$; D, $c_{F}=0.02$). Mean allele values (solid lines) and associated standard errors (shading) are calculated across all individuals within a population over 40000 generations across 40 replicate populations. Negative mean allele values indicate strategies of inbreeding avoidance or tendency for monandry, and positive values indicate strategies of inbreeding preference or tendency for polyandry.](figures/fixed_adapt.png)
